---
title: 'Part 2 : Chi-sqaure with lizards tails'
author: "Lauren Skube"
date: "12/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data 

```{r}
lizards <- read_csv(here("data" , "lizards.csv"))
```

Get counts of lizards by species 
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```
## Make a subset 

- Only contains UTST lizards 
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite), EAST (east tarbush site)
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B" , "W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site 

```{r}
utst %>% 
  count(site, tail)
```


```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

# create a table that shows counts and proportions by row
tail_proportions <- tail_counts %>%
  adorn_percentages() %>%
  janitor::adorn_pct_formatting(digits = 2) %>%
  janitor::adorn_ns()
  
```

## Convert to a contigency table 

- Use `column_to_rowname()`

```{r}
lizards_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square 

- Are tail breakage and site independent? If so, we would expect similar proportions of lizards with broken or whole tails across the three sites 

```{r}
lizard_x2 <- chisq.test(lizards_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```

There is sufficient evidence to reject the null hypothesis of independence and retail the alternate hypothesis that site and tail breakage are not independent. 

There is a significant association (i.e. non-idependence) between site and lizard tail breakage ($\chi$^2(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic` , *p* = `r lizard_tidy$p.value`)

There is a significant effect of site on tail breakage.


